---
title: "P8106 Data Science II Homework 3: Predicting Gas Milage"
author: "Sarah Forrest - sef2183"
date: "3/24/2023"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
--- 

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Data

In this exercise, we build a model to predict whether a given car gets high or low gas mileage based on a set of predictors from the dataset “auto.csv”. The dataset contains 392 observations. The response variable is `mpg_cat`, which indicates whether the miles per gallon of a car is high or low. 

```{r}
# read in data
auto = read.csv("data/auto.csv") 
```

Split the dataset into two parts: training data (70%) and test data (30%).

```{r}
set.seed(1)

# specify rows of training data (70% of the dataset)
trRows <- createDataPartition(auto$mpg_cat, 
                              p = .7,
                              list = F)

# training data
auto_train <- auto[trRows, ]
## matrix of predictors
x <- model.matrix(mpg_cat~.,auto)[trRows,-1]
## vector of response
y <- auto$mpg_cat[trRows]

# test data
auto_test <- auto[-trRows, ]
## matrix of predictors
x2 <- model.matrix(mpg_cat~.,auto)[-trRows,-1]
## vector of response
y2 <- auto$mpg_cat[-trRows]
```

# (a) Perform a logistic regression using the training data.

```{r}
set.seed(1)
```

Predictors in the model [APPEAR/DO NOT APPEAR] to be statistically significant? [If so, which ones?] 

**Confusion matrix using the test data with a probability threshold to determine class labels**

```{r}
```

The confusion matrix is telling [].

# (b) Train a multivariate adaptive regression spline (MARS) model using the training data.

```{r}
```

# (c) Perform LDA using the training data.

```{r}
```

## Plot of the linear discriminants in LDA

# (d) Prediction of the response variable.

```{r}
```

I would use the [] model to predict the response variable, `mpg_cat`. 

## Plot of ROC curve using the test data. 

```{r}
```

The AUC is []. The misclassification error rate is [].